---
title: "Process model results for NO2 instead of NOX"
author: "Robbie M Parks"
date: "10/06/2021"
output:
  html_document: 
    toc: yes
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Declare root directory, load folder locations, packages
```{r include=FALSE}
rm(list=ls())

# Declare root directory, load folder locations packages
project.folder <- paste0(print(here::here()),'/')
source(paste0(project.folder,'0_00_create_folder_structure.R'))
source(paste0(file.locations.folder,'file_locations.R'))
source(paste0(packages.folder,'packages_to_load.R'))
```

# Source functions to process results 
```{r include=FALSE}
source(paste0(model.processing.folder,'d_00_model_processing_functions.R'))
```

# Change output folder for reading model results to NO2 folder just for here 
```{r include=FALSE}
output.folder = output.no2.folder
```

# Load results
```{r include=FALSE}
# list of models and datasets
fmls = paste0('fml',c(1:116))
datasets = c('data_1yr','data_5yrs','data_10yrs')

# loop through models and datasets with various models
for(j in c(1:length(datasets))){
  for(i in c(101:116)){
    model_current <- paste0('mod','_',fmls[i],'_',datasets[j])
    if (file.exists(paste0(output.folder,model_current,'.rds'))) {
      assign(model_current, readRDS(paste0(output.folder,model_current,'.rds')))
    }
  }}
```

# process results for NO2
```{r include=FALSE}
dat.mean.exp <- data.frame()
for(j in c(1:length(datasets))){

    # STAN processing
    for(i in c(101,102,103,106:115)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STAN(fmls[i],datasets[j],'NO2')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
      }
    }

    # STAN ARM processing
    for(i in c(116)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STANARM(fmls[i],datasets[j],'NO2')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
      }
    }

}
dat.no2 <- dat.mean.exp
```

# process results for CO
```{r include=FALSE}
dat.mean.exp <- data.frame()
for(j in c(1:length(datasets))){

    # STAN processing
    for(i in c(101,102,104,106:115)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STAN(fmls[i],datasets[j],'CO')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
      }
    }
  
    # STAN ARM processing
    for(i in c(116)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STANARM(fmls[i],datasets[j],'CO')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
      }
    }
  
}
dat.co <- dat.mean.exp
```

# process results for EC
```{r include=FALSE}
dat.mean.exp <- data.frame()
for(j in c(1:length(datasets))){

    # STAN processing
    for(i in c(101,102,105,106:115)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STAN(fmls[i],datasets[j],'EC')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
       }
    }

      # STAN ARM processing
    for(i in c(116)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STANARM(fmls[i],datasets[j],'EC')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
       }
    }
}
dat.ec <- dat.mean.exp
```

# process results for O3
```{r include=FALSE}
dat.mean.exp <- data.frame()
for(j in c(1:length(datasets))){

    # STAN processing
    for(i in c(102)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STAN(fmls[i],datasets[j],'O3')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
       }
    }

}
dat.o3 <- dat.mean.exp
```

# process results for PM2.5
```{r include=FALSE}
dat.mean.exp <- data.frame()
for(j in c(1:length(datasets))){

    # STAN processing
    for(i in c(101:115)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STAN(fmls[i],datasets[j],'PM25')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
       }
    }
  
      # STAN ARM processing
    for(i in c(116)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STANARM(fmls[i],datasets[j],'PM25')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
       }
    }

}
dat.pm <- dat.mean.exp
```

# process results for Traffic Total
```{r include=FALSE}
dat.mean.exp <- data.frame()
for(j in c(1:length(datasets))){

    # STAN processing
    for(i in c(101:102,106:115)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STAN(fmls[i],datasets[j],'Traffic Total')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
       }
    }

}
dat.traffic.total <- dat.mean.exp
```

# process results for Traffic Average
```{r include=FALSE}
dat.mean.exp <- data.frame()
for(j in c(1:length(datasets))){

    # STAN processing
    for(i in c(101:102,106:115)){
      if (file.exists(paste0(output.folder,'mod_', fmls[i], '_', datasets[j],'.rds'))) {
        dat.temp = process_STAN(fmls[i],datasets[j],'Traffic Average')
        dat.mean.exp = rbind(dat.mean.exp, dat.temp)
       }
    }

}
dat.traffic.average <- dat.mean.exp
```

# combine results for all pollutants and save
```{r include=FALSE}

dat.all = rbind(dat.pm, dat.no2, dat.co, dat.ec, dat.o3, dat.traffic.total,dat.traffic.average) 
  
# reorder formulas
dat.all$mod = ordered(dat.all$mod, levels = paste0('fml', c(101:116)))
dat.all$pollutant = ordered(dat.all$pollutant, levels = c('Traffic Total','Traffic Average','NO2','CO','EC', 'PM25', 'O3'))

dir.output = paste0(figures.folder,'figure2/')
ifelse(!dir.exists(dir.output), dir.create(dir.output, recursive=TRUE), FALSE)
write.csv(dat.all, paste0(dir.output,'multi_pollutant_output_no2.csv'))
print(dat.all)
```
