---
title: "Prepare participant data"
author: "Robbie M Parks"
date: "16/06/2022"
output:
  html_document: 
    toc: yes
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Declare root directory, load folder locations, packages
```{r include=FALSE}
rm(list=ls())

# Declare root directory, load folder locations packages
project.folder <- paste0(print(here::here()),'/')
source(paste0(project.folder,'0_00_create_folder_structure.R'))
source(paste0(file.locations.folder,'file_locations.R'))
source(paste0(packages.folder,'packages_to_load.R'))
```

# Reading in patient and parish data
```{r}
# Load project with patient data
load(paste0(raw_data_files, "ALS_desc_1989_2013_1989_5_2k.RData"))
patient_data <- ALS_desc_1989_2013_1989_5_2k 

summary(patient_data)

# load parish ses data
parish_ses <- read_sas(paste0(raw_data_files, "sogn_udd_2009b_1R.sas7bdat"), NULL)
```

# Defining variables for analysis
## Converting categorical variables to factors and labeling the factors based on dictionary
```{r include=FALSE}
patient_data = patient_data %>%
  mutate(diagyr=year(diagdat))
patient_data$sex <- as.factor(patient_data$sex)
patient_data$social <-factor(patient_data$social)
patient_data$status <-factor(patient_data$status)
patient_data$pattype<-factor(patient_data$pattype)
patient_data$civst <- factor(patient_data$civst)
patient_data$ck <- factor(patient_data$ck)
patient_data$aegtesocial <- as.factor(patient_data$aegtesocial)
patient_data$diag<-as.factor(patient_data$diag)
patient_data$diagyr <- as.factor(patient_data$diagyr)
```

## Coding the NA values
```{r include=FALSE}
#namaker converts the empty strings to NA
namaker<- function(a){
  ifelse(a==""|a=="-"|a=="----"|a==" -",is.na(a)==TRUE,a)
    }

#converting the missing codes to NA
patient_data$social <- ifelse(patient_data$social==9,NA,patient_data$social)
patient_data$aegtesocial <- ifelse(patient_data$aegtesocial==9,NA,patient_data$aegtesocial)
```

## Calculating age at diagnosis
```{r}
patient_data$age_diag = time_length(difftime(patient_data$diagdat, patient_data$bdate), unit = "years")
```

## Calculating place of birth 
```{r}
patient_data <- patient_data %>%
mutate(place_birth = case_when(geokode >= 52 & geokode <= 62 ~ "foreign",
                                 geokode == 99 | geokode == 10 ~ "unknown",
                                 geokode > 1 & geokode <= 4 ~ "other_cities",
                                 geokode == 1 ~ "copenhagen",
                                 geokode >= 5 & geokode <= 9 ~ "other_dk",
                                 geokode == 51 ~ "greenland"),
         place_birth = as.factor(place_birth))

summary(patient_data)
```

# Creating a variable to categorize participants based on place of residence at time of diagnosis based on zipcodes
# Based on https://postal-codes.cybo.com/denmark/
# Based on Denmark cities larger than 50,000 (https://worldpopulationreview.com/countries/cities/denmark)
```{r}
patient_data <- patient_data %>%
 mutate(residence = if_else(postnr == 3950 |
                            postnr == 3985 |
                            postnr == 3952 |
                            postnr == 3910 |
                            postnr == 3912 |
                            postnr == 3921 |
                            postnr == 3905 |
                            postnr == 3971 |
                            postnr == 3951 |
                            postnr == 3911 |
                            postnr == 3913 |
                            postnr == 3919 |
                            postnr == 3984 |
                            postnr == 3980 |
                            postnr == 3930 |
                            postnr == 3982 |
                            postnr == 3923 |
                            postnr == 3940 |
                            postnr == 3964 |
                            postnr == 3953 |
                            postnr == 3992 |
                            postnr == 3962, "Greenland",
                    if_else(postnr >= 1050 & postnr <= 4050, "Greater Copenhagen", # Copenhagen, Capital Region of Denmark
                    if_else(postnr >= 8000 & postnr <= 8381, "Big cities of Denmark>10,000", # Aarhus
                    if_else(postnr >= 5000 & postnr <= 5320, "Big cities of Denmark>10,000", # Odense
                    if_else(postnr >= 9000 & postnr <= 9220, "Big cities of Denmark>10,000", # Aalborg
                    if_else(postnr >= 6700 & postnr <= 6715, "Big cities of Denmark>10,000", # Esbjerg
                    if_else(postnr >= 8900 & postnr < 8960,  "Big cities of Denmark>10,000", # Randers
                    if_else(postnr == 6000, "Big cities of Denmark>10,000",                  # Kolding
                    if_else(postnr == 7100 | postnr == 7120, "Big cities of Denmark>10,000", # Vejle
                    if_else(postnr == 8700, "Big cities of Denmark>10,000", # Horsens
                    if_else(postnr == 8830, "Big cities of Denmark>10,000", # Randers
                            "Rest of Denmark")))))))))))) %>%
  mutate(residence = replace_na(residence, "Unknown")) %>%
  mutate(residence = as.factor(residence))

table(patient_data$residence)

summary(patient_data)

patient_data_residence_na = patient_data %>% filter(residence=='Unknown')

print(patient_data_residence_na)
```

# (previous code now unused)
#Creating a variable to categorize participants based on place of residence at time of diagnosis based on zipcodes
## Alternate residence 
```{r}
# patient_data$residence <- ifelse(patient_data$postnr>=1000 & patient_data$postnr<2500, 1, ifelse(patient_data$postnr==8722|patient_data$postnr==4690|patient_data$postnr==2800|patient_data$postnr==8300|patient_data$postnr==2791|patient_data$postnr==9500|patient_data$postnr==2640|patient_data$postnr==7600|patient_data$postnr==3060|patient_data$postnr==9700|patient_data$postnr==4760|patient_data$postnr==3300|patient_data$postnr==6800|patient_data$postnr==7700|patient_data$postnr==3500|patient_data$postnr==2625|patient_data$postnr==2665|patient_data$postnr==8660|patient_data$postnr==5500|patient_data$postnr==3700|patient_data$postnr==4900|patient_data$postnr==8500|patient_data$postnr==3600|patient_data$postnr==7430|patient_data$postnr==2680|patient_data$postnr==4220|patient_data$postnr==5800|patient_data$postnr==4400|patient_data$postnr==6200|patient_data$postnr==4800|patient_data$postnr==3520|patient_data$postnr==3460|patient_data$postnr==2635|patient_data$postnr==3660|patient_data$postnr==4100|patient_data$postnr==4129|patient_data$postnr==2600|patient_data$postnr==7800|patient_data$postnr==6100|patient_data$postnr==9400|patient_data$postnr==9900|patient_data$postnr==4300|patient_data$postnr==9800|patient_data$postnr==6400|patient_data$postnr==5700|patient_data$postnr==2610|patient_data$postnr==2620|patient_data$postnr==3400|patient_data$postnr==2630|patient_data$postnr==4200|patient_data$postnr==7500|patient_data$postnr==4600|patient_data$postnr==8800|patient_data$postnr==3000|patient_data$postnr==2970|patient_data$postnr==7000|patient_data$postnr==7007|patient_data$postnr==7029|patient_data$postnr==2720|patient_data$postnr==2750|patient_data$postnr==2920|patient_data$postnr==4700|patient_data$postnr==8600|patient_data$postnr==4000|patient_data$postnr==7400|patient_data$postnr==7429|patient_data$postnr==7429|patient_data$postnr==2670|patient_data$postnr==2650|patient_data$postnr==8700|patient_data$postnr==7100|patient_data$postnr==7120|patient_data$postnr==6000|patient_data$postnr==8960|patient_data$postnr==8900|patient_data$postnr==6715|patient_data$postnr==6700|patient_data$postnr==9220|patient_data$postnr==9000|patient_data$postnr==5270|patient_data$postnr==5000|patient_data$postnr==8000|patient_data$postnr==8210, 2,ifelse(patient_data$postnr>=3900 & patient_data$postnr<3993,3,ifelse(is.na(patient_data$postnr),5,4))) ) 
# patient_data$residence <- ifelse(is.na(patient_data$postnr),5,patient_data$residence)
# patient_data$residence<- factor(patient_data$residence,levels=c('1','2','3','4','5'), labels=c("Greater Copenhagen", "Big cities of Denmark>10,000","Greenland", "Rest of Denmark","Unknown"))
# 
# table(patient_data$residence)
# 
# summary(patient_data)
```

# Save output
```{r}
saveRDS(patient_data, paste0(raw_data_files,"patient_data_uniform_complete.rds"))
```
