---
title: "Creating 1-, 5- and 10-year averages for windows of exposure"
author: "Robbie Parks"
date: "16/06/20221"
output:
  html_document: 
    toc: yes
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Declare root directory, load folder locations, packages
```{r include=FALSE}
rm(list=ls())

# Declare root directory, load folder locations packages
project.folder <- paste0(print(here::here()),'/')
source(paste0(project.folder,'0_00_create_folder_structure.R'))
source(paste0(file.locations.folder,'file_locations.R'))
source(paste0(packages.folder,'packages_to_load.R'))
```

# Load merged dataset (large dataset so be careful)
```{r}
full_dataset <- readRDS(paste0(robbie_datasets, "joined_datasets.rds"))
summary(full_dataset)
```

# Nest data by cases so we can easily create windows of exposure for each case
```{r}
# A data set made up of sub-datasets. Each data subset corresponds to a patient and his/her controls
nested_data <- full_dataset %>% 
  mutate(nr = as.factor(nr),
         lobnr = as.factor(lobnr),
         ck = as.factor(ck)) %>%
  group_by(nr) %>%
  nest()

# take a look in a single case and corresponding controls 
test <- as.data.frame(nested_data$data[[1]]) %>%
 dplyr::filter(date_exp <= year_gap & date_exp >= one_year)

summary(test)
```

# For each case, select years within exposure window of 1, 5, and 10 years pre-hospitalization with a 1-year gap between exposure and first hospitalization
```{r}
# filter data for the 1, 5 and 10 year periods
filtered_data <- nested_data %>%
  mutate(one_year = purrr::map(data, ~ dplyr::filter(.x, date_exp <= year_gap & date_exp > one_year)),
         five_yr = purrr::map(data, ~ dplyr::filter(.x, date_exp <= year_gap & date_exp > five_years)),
         ten_yr = purrr::map(data, ~ dplyr::filter(.x, date_exp <= year_gap & date_exp > ten_years)))
 
# explore a random dataset within the list of datasets to make sure it looks right
explore <- filtered_data$one_year[[1]]
summary(explore)

# save list of dataframes containing the datasets for each time frame
#save(filtered_data, file = paste0(robbie_datasets, "data_frames_nested.RData"))
```

# Re-load list of dataframes once already run
```{r, include=FALSE}
#load(file = paste0(robbie_datasets, "data_frames_nested.RData")) # this is also a large dataset
```

# Explore missing data in the 1-year time window exposure and save the dataset
```{r}
# select the 1-year time window dataset
one_year <- filtered_data %>%
  dplyr::select(-data, -five_yr, -ten_yr) %>%
  unnest(., cols = one_year)

# check how many cases and controls total we have in this data set
case_num <- one_year %>%
  dplyr::filter(ck == 1) %>% # selecting only cases
  count(nr) # count number of unique "nr" (cases)

control_num <- one_year %>%
  dplyr::filter(ck == 0) %>% # selecting only controls
  group_by(nr) %>% 
  count(lobnr)  

# save one year time window data
#write_csv(one_year, paste0(robbie_datasets, "one_yr_data.csv"))

# check how many cases/controls are missing months
exp_missing <- one_year %>%
 group_by(nr, lobnr) %>%
  summarise(counts = n()) # lowest number of months measure for a given case or control is 1

exp_missing %>%
  tally(counts<12) %>%
  summarize(n=sum(n))

# heat map to visualize missingness (white means missing)
exp_missing %>% ggplot(aes(as.numeric(nr), lobnr, fill= counts)) + 
  geom_tile() +
  theme_classic() +
  scale_fill_gradient('counts', limits=c(1, 12), breaks = c(1, 3, 6, 9, 12),  low = "red", high = "darkblue")
```

# Explore missing data in the 5-year time window exposure and save the dataset
```{r}
# select the 5-years time window dataset
five_years <- filtered_data %>%
  dplyr::select(-data, -one_year, -ten_yr) %>%
  unnest(., cols = five_yr)

# check how many cases and controls total we have in this data set
case_num_5 <- five_years %>%
 dplyr::filter(ck == 1) %>% # selecting only cases
  count(nr) # count number of unique "nr" (cases)

# check how many controls total we have in this dataset
control_num_5 <- five_years %>%
  dplyr::filter(ck == 0) %>% # selecting controls
  group_by(nr) %>% 
   count(lobnr)

# save the 5-years time window dataframe
#write_csv(five_years, paste0(robbie_datasets, "five_yrs_data.csv"))

# check how many cases/controls are missing months
exp_missing_5yrs <- five_years %>%
  group_by(nr, lobnr) %>%
  summarise(counts = n()) # minimum number of available months is 1

exp_missing_5yrs %>%
  tally(counts<12) %>%
  summarize(n=sum(n))

# heat map to visualize missingness (white means missing)
exp_missing_5yrs %>% ggplot(aes(as.numeric(nr), lobnr, fill= counts)) + 
  geom_tile() +
  theme_classic() +
  scale_fill_gradient('counts', limits=c(1, 60), breaks = c(1, 12, 24, 36, 48, 60),  low = "red", high = "darkblue") 
```

# Explore missing data in the 10-year time window exposure and save the dataset
```{r}
# select the ten-years time window dataset
ten_years <- filtered_data %>%
  dplyr::select(-data, -one_year, -five_yr) %>%
  unnest(., cols = ten_yr)

# check how many cases total we have in this dataset
case_num_10 <- ten_years %>%
 dplyr::filter(ck == 1) %>% # selecting only cases
  count(nr) # count number of unique "nr" (cases)

# check how many controls total we have in this dataset
control_num_10 <- ten_years %>%
  dplyr::filter(ck == 0) %>% # selecting controls
  group_by(nr) %>% 
   count(lobnr)

# save the 10-years time window dataframe
#write_csv(ten_years, paste0(robbie_datasets, "ten_yrs_data.csv"))

# check how many cases/controls are missing months
exp_missing_10yrs <- ten_years %>%
  group_by(nr, lobnr) %>%
  summarise(counts = n()) # minimum number of available months is 1

exp_missing_10yrs %>%
  tally(counts<12) %>%
  summarize(n=sum(n))

# heat map to visualize missingness (white means missing)
exp_missing_10yrs %>% ggplot(aes(as.numeric(nr), lobnr, fill= counts)) + geom_tile() +
  theme_classic() +
  scale_fill_gradient('counts', limits=c(1, 120), breaks = c(1, 24, 48, 72, 96, 120),  low = "red", high = "darkblue") 
```

~~~~~~~~~~~~~~~~~~~~~
Notes: 

For the 1-year exposure window, limit cases and controls to those we have at least 75% of the pollutant measurements. That is, at least **9 out of the 12 months** need to have a pollutant measurement. Furthermore, within those 9 measurements we need to have at least one measure for each of the year seasons (Denmark has 4 seasons) 

For the 5-years and 10-years exposure window, limit cases and controls to those we have at least 50% of the PM2.5 measurements. That is, for the 5-year time window at least **30 out of the 60 months** should have a measurement. For the 10-year window, each case/control should have at least 60 months with pollutant measurements out of the 120 months. 

For all time windows, each case needs to have at least one control that fulfills the requirements described in the above text. 

# Filtered cases/controls that satisfy the above requirements 

# 1-year exposure windows

****** Check total number of cases *****
```{r}
filter_1yr <- one_year %>%
  group_by(nr, lobnr) %>%
  dplyr::mutate(months_meas = n()) %>%
  ungroup() %>%
  dplyr::filter(months_meas >= 9) %>% # select those with at least 9 months 
  mutate(month_exp = month(date_exp),
         season = if_else(month_exp >=3 & month_exp <= 5, "spring", 
                  if_else(month_exp >=6 & month_exp <=8, "summer",
                  if_else(month_exp >=9 & month_exp <= 11, "fall",
                  if_else(month_exp == 12 | month_exp <= 2, "winter", "x"))))) %>%  # season ID
  group_by(nr, lobnr) %>%
  mutate(season_num = length(unique(season))) %>% #XX entries are missing measurements in a season
ungroup() %>%
  dplyr::filter(season_num >= 4) %>% # filter for entries that have at least 1 measurement in each season
group_by(nr) %>%
  dplyr::mutate(case_control = length(unique(ck))) %>% # ID to check that each case has at least one control and that each control has a case 
  dplyr::filter(case_control >= 2) %>% # filter cases that have  at least  one  control  and controls that have a  case
  ungroup() %>%
  dplyr::select(-month_exp, -months_meas, -season, -season_num, -case_control) # remove IDs
         
summary(filter_1yr)

# final number of cases
length(unique(filter_1yr$nr))

#final number of controls
filter_1yr %>%
  dplyr::filter(ck == 0) %>% #select controls only
  group_by(nr) %>%
 count(lobnr) %>%
  nrow()
    
# save data 
write_csv(filter_1yr, paste0(robbie_datasets, "one_yr_filtered_data.csv"))
```

# 5-years exposure windows

****** Check total number of cases *****
```{r}
filter_5yrs <- five_years %>%
  group_by(nr, lobnr) %>%
  dplyr::mutate(months_meas = n()) %>% #ID to check mounts measure
  ungroup() %>%
  dplyr::filter(months_meas >= 30) %>% # select those with at least 30 months 
group_by(nr) %>%
  dplyr::mutate(case_control = length(unique(ck))) %>% # ID to check that each case has at least one control and that each control has a case 
  dplyr::filter(case_control >= 2) %>%
  ungroup() %>%
  dplyr::select(-case_control, -months_meas) # remove IDs
         
summary(filter_5yrs)

# final number of cases
length(unique(filter_5yrs$nr))

#final number of controls
filter_5yrs %>%
  dplyr::filter(ck == 0) %>% #select controls only
  group_by(nr) %>%
 count(lobnr) %>%
  nrow()
    
# save data 
write_csv(filter_5yrs, paste0(robbie_datasets, "five_yrs_filtered_data.csv"))
```

# 10-years exposure window 

****** Check total number of cases *****
```{r}
filter_10yrs <- ten_years %>%
  group_by(nr, lobnr) %>%
  dplyr::mutate(months_meas = n()) %>% #ID to check months measure
  ungroup() %>%
  dplyr::filter(months_meas >= 60) %>% # select those with at least 60 months 
group_by(nr) %>%
  dplyr::mutate(case_control = length(unique(ck))) %>% # ID to check that each case has at least one control and that each control has a case 
  dplyr::filter(case_control >= 2) %>%
  ungroup() %>%
  dplyr::select(-case_control, -months_meas) # remove ID
         
summary(filter_10yrs)

# final number of cases
length(unique(filter_10yrs$nr))

#final number of controls
filter_10yrs %>%
 dplyr::filter(ck == 0) %>% #select controls only
  group_by(nr) %>%
 count(lobnr) %>%
  nrow()
    
# save data 
write_csv(filter_10yrs, paste0(robbie_datasets, "ten_yrs_filtered_data.csv"))
```
