---
title: "Find overall averages of exposure windows"
author: "Robbie Parks"
date: "16/06/2022"
output:
  html_document: 
    toc: yes
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Declare root directory, load folder locations, packages
```{r include=FALSE}
rm(list=ls())

# Declare root directory, load folder locations packages
project.folder <- paste0(print(here::here()),'/')
source(paste0(project.folder,'0_00_create_folder_structure.R'))
source(paste0(file.locations.folder,'file_locations.R'))
source(paste0(packages.folder,'packages_to_load.R'))
```

For all dataframes, we created averages for given exposure period and added the family income variable and removed the CO-urb variable that we forgot to remove previously. Family income (family_SES) is base in the highest social when married and in the individual social if single. 

- We also categorized the place of birth into these categories:
1) Greater Copenhagen
2) Other cities: includes Aarhus, Odense, and other DK cities
3) Other DK: includes everything else other than major cities
4) Foreign: includes non danish regions
5) Unknown: unknown DK or unknown
6) Greenland: includes Greenland

- The county of residence was also categorized from zip code to the following five categories: 
1) Greater Copenhagen
2) Other cities: includes Aarhus, Odense, and other DK cities
3) Other_DK: includes everything else other than major cities
4) Foreign: includes non danish regions
5) Unknown: unknown DK or unknown
6) Greenland: includes Greenland

# 1-year exposure averages
```{r}
avg_1yr <- read_csv(paste0(robbie_datasets, "one_yr_filtered_data.csv")) %>%
  group_by(nr, lobnr) %>%
  mutate(PM25_1yr = mean(PM25_str),
         CO_1yr = mean(CO_str),
         NO2_1yr = mean(NO2_str),
         NOX_1yr = mean(NOX_str),
         O3_1yr = mean(O3_str),
         PM10_1yr = mean(PM10_str),
         EC_1yr = mean (EC_str),
         family_SES = as.factor(family_SES)) %>%
  ungroup() %>%
   dplyr::select(-year_gap, -one_year, -five_years, -ten_years, -PM25_str, -CO_str, -EC_str, -NOX_str, -O3_str, -PM10_str, -NO2_str, -date_exp)%>%
  unique() 

summary(avg_1yr)

# check that we have the correct number of cases 
check_cases <- avg_1yr %>%
  group_by(nr) %>%
  summarise()

print(nrow(check_cases))

# save data file 
write_csv(avg_1yr, paste0(robbie_datasets, "one_yr_avg_data.csv"))
```

# 5-year exposure averages
```{r}
avg_5yr <- read_csv(paste0(robbie_datasets, "five_yrs_filtered_data.csv")) %>%
  group_by(nr, lobnr) %>%
  mutate(PM25_5yr = mean(PM25_str),
         CO_5yr = mean(CO_str),
         NO2_5yr = mean(NO2_str),
         NOX_5yr = mean(NOX_str),
         O3_5yr = mean(O3_str),
         PM10_5yr = mean(PM10_str),
         EC_5yr = mean (EC_str),
         family_SES = as.factor(family_SES)) %>%
  ungroup() %>%
   dplyr::select(-year_gap, -one_year, -five_years, -ten_years, -PM25_str, -CO_str, -EC_str, -NOX_str, -O3_str, -PM10_str, -NO2_str, -date_exp)%>%
  unique()
    
summary(avg_5yr)

# check that we have the correct number of cases 
check_cases <- avg_5yr %>%
  group_by(nr) %>%
  summarise()

print(nrow(check_cases))

# save data file 
write_csv(avg_5yr, paste0(robbie_datasets, "five_yrs_avg_data.csv"))
```

# 10-year exposure averages
```{r}

avg_10yr <- read_csv(paste0(robbie_datasets, "ten_yrs_filtered_data.csv")) %>%
  group_by(nr, lobnr) %>%
  mutate(PM25_10yr = mean(PM25_str),
         CO_10yr = mean(CO_str),
         NO2_10yr = mean(NO2_str),
         NOX_10yr = mean(NOX_str),
         O3_10yr = mean(O3_str),
         PM10_10yr = mean(PM10_str),
         EC_10yr = mean (EC_str),
         family_SES = as.factor(family_SES)) %>%
  ungroup() %>%
   dplyr::select(-year_gap, -one_year, -five_years, -ten_years, -PM25_str, -CO_str, -EC_str, -NOX_str, -O3_str, -PM10_str, -NO2_str, -date_exp)%>%
  unique()

summary(avg_10yr)

# check that we have the correct number of cases 
check_cases <- avg_10yr %>%
  group_by(nr) %>%
  summarise()

print(nrow(check_cases))

# save data file 
write_csv(avg_10yr, paste0(robbie_datasets, "ten_yrs_avg_data.csv"))
```
